---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# hcorp: Music corpora for harmonic analysis

[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)
[![Travis build status](https://travis-ci.org/pmcharrison/hcorp.svg?branch=master)](https://travis-ci.org/pmcharrison/hcorp)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/github/pmcharrison/hcorp?branch=master&svg=true)](https://ci.appveyor.com/project/pmcharrison/hcorp)
[![Coverage status](https://coveralls.io/repos/github/pmcharrison/hcorp/badge.svg)](https://coveralls.io/r/pmcharrison/hcorp?branch=master)
[![DOI](https://zenodo.org/badge/DOI/10.5281/zenodo.2545754.svg)](https://doi.org/10.5281/zenodo.2545754)

This R package provides several datasets of chord sequences.
These datasets are expressly for research purposes only.

- `bach_chorales_1`: 370 chorales by J. S. bach from 
[KernScores](http://kern.humdrum.org/)
- `classical_1`: 1,022 classical pieces compiled from
[KernScores](http://kern.humdrum.org/),
converted to chord sequences using the algorithm of 
Pardo & Birmingham (2002).
- `popular_1`: 739 pieces from the McGill Billboard corpus (Burgoyne, 2011),
converted from chord symbols to pitch-class sets by Harrison & Pearce (2018).
- `jazz_1`: 1,186 pieces from the iRB corpus (Broze & Shanahan, 2013),
converted from chord symbols to pitch-class sets by Harrison & Pearce (2018).

For more details, see the package's documentation (e.g. `?classical_1`).

## Installation

You can install the current version of `hcorp` from Github
by entering the following commands into R:

``` r
if (!require(devtools)) install.packages("devtools")
devtools::install_github("hcorp")
```

## Example usage

The `hcorp` package is best used in tandem with the `hrep` package.
The `hrep` package provides the underlying representations for the corpora
in `hcorp`,
as well as methods for manipulating and visualising them.

You can load these packages into the global namespace as follows:

```{r}
library(hcorp)
library(hrep)
library(magrittr) # Provides the pipe operator, %>%
```

The `hrep` package currently contains three corpora:

```{r}
classical_1

popular_1

jazz_1
```

Internally, a corpus is a list of encoded vectors.

```{r}
classical_1[1:3] %>% as.list
```

Encoded vectors are objects of class `coded_vec`.

```{r}
classical_1[[1]] %>% class
```

Internally, encoded vectors are sequences of integers.
This is good for memory efficiency,
and useful for certain modelling approaches.

```{r}
classical_1[[1]] %>% as.integer %>% head
```

These vectors can be decoded with the function `decode`.

```{r}
classical_1[[1]][1:3] %>% decode

classical_1[[1]][1:3] %>% decode %>% as.list
```

Corpora and sequences can optionally store metadata.

```{r}
metadata(classical_1)

metadata(classical_1[[1]])
```

Corpora and sequences can be subsetted and combined like lists.

```{r}
classical_1[1:3]

classical_1[[1]]

classical_1[[1]][1:3]

c(classical_1[1:3],
  popular_1[1:3])
```

## References

Broze, Y., & Shanahan, D. (2013). Diachronic changes in jazz harmony: A cognitive perspective. Music Perception, 31(1), 32–45. https://doi.org/10.1525/rep.2008.104.1.92

Harrison, P. M. C., & Pearce, M. T. (2018). An energy-based generative sequence model for testing sensory theories of Western harmony. In Proceedings of the 19th International Society for Music Information Retrieval Conference (pp. 160–167). Paris, France.

Pardo, B., & Birmingham, W. P. (2002). Algorithms for chordal analysis. Computer Music Journal, 26(2), 27–49. https://doi.org/10.1162/014892602760137167
